---
title: 'Create Map - Earthquake Analysis'
author: "Alaleh Ahmadian"
date: "May 8, 2018"
output:
  prettydoc::html_pretty:
    highlight: github
    theme: HPSTR
  rmarkdown::html_document: default
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```


## Libraries:

```{r , message=FALSE, warning=FALSE}

library(readr)
library(dplyr)
library(highcharter)
library(ggplot2)
library(tidyr)
library(plotly)
library(readtext)
library(readr)
library(stringr)
library(magrittr)
library(widgetframe)
library(rworldmap)
library(ggthemes)
library(ggmap)
library(gganimate)
library(animation)
library(gapminder)
library(sp)
library(mapdata)
library(maps)
library(rgeos)
library(maptools)

```


## Data:

changes daster.txt to disaster.csv to be able to read the data better 

```{r , message=FALSE, warning=FALSE}

hist_data = readRDS("C:/Users/Alaleh/Desktop/Term8/Data_Analysis/Data/earthquake/historical_web_data_26112015.rds")
iran_earthquake = readRDS("C:/Users/Alaleh/Desktop/Term8/Data_Analysis/Data/earthquake/iran_earthquake.rds")
disaster = read.csv(file="C:/Users/Alaleh/Desktop/Term8/Data_Analysis/HW_sol/disaster.csv", header=T,  row.names=NULL)
colnames(disaster) <- c(colnames(disaster)[-1],"x")
disaster$x <- NULL
worldwide = read.csv(file="C:/Users/Alaleh/Desktop/Term8/Data_Analysis/Data/earthquake/worldwide.csv", header=T,  row.names=NULL)


```



## 1

We omit data from Afghanistan, Azerbaijan and Pakistan because they're not in iran and also have too many outliers which can make the plot unreadable (to see the full plot with these countries you just have to comment the 3rd line)

```{r , message=FALSE, warning=FALSE}

colors <- c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951')

foreigns = c("Afghanistan","Azerbaijan","Pakistan")
data_3D = hist_data[!(hist_data$Province %in% foreigns),]

plot_ly(data_3D, x = ~Latitude, y = ~Longitude, z = ~Depth, color = ~City, size = ~Magnitude, colors = colors,
             marker = list(symbol = 'circle', sizemode = 'diameter'), sizes = c(1.5,40),
             text = ~paste('City:', City, '<br>Magnitude:', Magnitude, '<br>Province:', Province)) %>% layout(title = 'Place and depth of earthquake based on its magnitude',
         scene = list(xaxis = list(title = 'Latitude',
                      gridcolor = 'rgb(255, 255, 255)',
                      zerolinewidth = 1,ticklen = 5,
                      gridwidth = 2),
               yaxis = list(title = 'Longitude',
                      gridcolor = 'rgb(255, 255, 255)',
                      zerolinewidth = 1,
                      ticklen = 5,
                      gridwith = 2),
               zaxis = list(title = 'Death',
                            gridcolor = 'rgb(255, 255, 255)',
                            zerolinewidth = 1,
                            ticklen = 5,
                            gridwith = 2)),
         paper_bgcolor = 'rgb(243, 243, 243)',
         plot_bgcolor = 'rgb(243, 243, 243)')



```





## 2


```{r , message=FALSE, warning=FALSE}

sitsunami = disaster

sitsunami$DAY[is.na(sitsunami$DAY)] = 1
sitsunami$MONTH[is.na(sitsunami$MONTH)] = 1
sitsunami$HOUR[is.na(sitsunami$HOUR)] = 1
sitsunami$MINUTE[is.na(sitsunami$MINUTE)] = 1
sitsunami$SECOND[is.na(sitsunami$SECOND)] = 1
sitsunami$YEAR[is.na(sitsunami$YEAR)] = 1
sitsunami$date <- as.Date(with(sitsunami, paste(YEAR, MONTH, DAY,sep="-")), "%Y-%m-%d")

sitsunami = sitsunami %>% filter(FLAG_TSUNAMI=="Tsu") %>%  rename(year=YEAR, month=MONTH, day=DAY, lat = LATITUDE,lon = LONGITUDE, z = INTENSITY,name = LOCATION_NAME) %>% select(lat, lon, z, name, date, year, month, day) %>% na.omit()

sitsunami$lat = as.numeric(as.character(sitsunami$lat))
sitsunami$lon = as.numeric(as.character(sitsunami$lon))

sitsunami %>% arrange(year) -> sitsunami

world <- ggplot() +
  borders("world", colour = "gray95", fill = "gray80") +
  theme_map() 

map <- world + geom_point(data = sitsunami, aes(x = lon, y = lat, size = z, frame = date, cumulative = TRUE),colour = 'purple', alpha = .5) + scale_size_continuous(range = c(1, 8), breaks = c(250, 500, 750, 1000)) + labs(size = 'Followers')  + coord_fixed(1.3)  + theme_map()

gganimate(map, interval = .07, ani.width=800, ani.height=600, "tsunamiss.gif", title_frame = T)


```




## 3

```{r , message=FALSE, warning=FALSE}

si_earthquakes = iran_earthquake %>% rename(lat = Lat,lon = Long, z = Mag,name = OriginTime) %>% select(lat, lon, z, name) %>% na.omit()

IRMap = get_map(geocode("Iran"),zoom = 5)

ggmap(IRMap) + geom_density2d(data = si_earthquakes, aes(x = lon, y = lat)) + stat_density2d(data = si_earthquakes, aes(x = lon, y = lat, fill = ..level.., alpha = ..level..), size = 0.01, bins = 16, geom = 'polygon') + scale_fill_gradient(low = "blue", high = "red") + scale_alpha(range = c(0.00, 0.25), guide = FALSE)  + theme(legend.position = "none", axis.title = element_blank(), text = element_text(size = 12))


```






## 4

Assuming Poisson distribution, a common choice of distribution for count data, the four earthquakes in ten years lead to maximum likelihood estimate of rate of earthquakes per year

??_hat = (???k_i) / n = 4/10

where k_i is the number of earthquakes in the i-th year. This leads to estimating probability of any earthquake in a single year as

P = Pr(K???1) = 1 ??? Pr(K=0) = 1 ??? (e^(?????)) ??^k / k!

Assuming independence between years, the probability of any earthquake in five years is equal to

1 ??? (Pr(K = 0))^5

(Pr(K = 0))^5 == no earthquake in 5 years

We want the opposite of the above 

Notice that this result is consistent with much simpler approach: observe that, according to your data, probability of observing a year with any earthquake is 3/10, so probability of observing at least one such year in five is 1???(1???3/10)^5

We could also use the gutenberg-richter law but the data didn't suffice for a good answer

```{r , message=FALSE, warning=FALSE}

iran_yearwise_eq = iran_earthquake %>% filter (Mag>=7)
iran_yearwise_eq$OriginTime = substr(iran_yearwise_eq$OriginTime,0,4)
iran_yearwise_eq %>% group_by(OriginTime) %>% summarise(count = n()) -> count_eq
count_eq

landa_hat = sum(count_eq$count)/10
P = 1 - exp(-landa_hat)
ans = 1 - (1-P)^5
ans

hchart(count_eq ,type = "column", title="Number of earthquakes in each year", hcaes(x = OriginTime, y=count))  %>% hc_add_theme(hc_theme_google())

```


## 5



```{r , message=FALSE, warning=FALSE}

world_ts_data = disaster %>% select(year = YEAR, country = COUNTRY, killed = DEATHS) %>% na.omit()  %>% group_by(country) %>% summarise(allDied = sum(as.integer(killed)), meanDead = mean(as.integer(killed)))

get_data_from_map(download_map_data("custom/world-eckert3")) %>% select(code = `hc-key`, country = name) %>% mutate(country = toupper(country)) -> map

world_ts_data$country = as.character(world_ts_data$country)
world_ts_data[world_ts_data == "USA"] = "UNITED STATES OF AMERICA"

all_map = left_join(map, world_ts_data) 

hcmap("custom/world-eckert3", 
      data = all_map, value = "meanDead", 
      joinBy = c("hc-key", "code"), name = "Mean Deaths", 
      borderColor = "blue", borderWidth = 0.2) %>% 
  hc_mapNavigation(enabled = TRUE) %>% hc_add_theme(hc_theme_google())

hcmap("custom/world-eckert3", 
      data = all_map, value = "allDied", 
      joinBy = c("hc-key", "code"), name = "Total Deaths", 
      borderColor = "blue", borderWidth = 0.2) %>% 
  hc_mapNavigation(enabled = TRUE) %>% hc_add_theme(hc_theme_google())

```



## 6


```{r , message=FALSE, warning=FALSE}

pred_data = disaster

pred_data = pred_data %>%  rename(lat = LATITUDE,lon = LONGITUDE, z = INTENSITY, depth = FOCAL_DEPTH, killed = TOTAL_DEATHS) %>% select(lat, lon, z, depth, killed)

pred_data$lat = as.numeric(as.character(pred_data$lat))
pred_data$lon = as.numeric(as.character(pred_data$lon))

fit = lm(killed~. , data = pred_data)
summary(fit)

fit2 = lm(killed ~ z + I(lat)^2 + I(lon)^2 + I(depth)^3 , data = pred_data)
summary(fit2)

```

We see that both predictions have a high error so we can't get a good prediction


## 7

```{r , message=FALSE, warning=FALSE}

worldwide %>% select(time, mag, place) %>% na.omit() %>% group_by(place,time) %>% filter(n()>1) -> time_place

time_place$place = as.character(time_place$place)
time_place$date = as.Date(str_sub(time_place$time,1,10),format="%Y-%m-%d")
time_place$time = as.character(str_sub(time_place$time,12,19))

foreData = cbind(time_place$mag[-length(time_place)], time_place$mag[-1])
foreData = data.frame(foreData[c(TRUE,FALSE),])
names(foreData) = c("foreshock","mainquake")

fit = lm(mainquake~foreshock, foreData)
summary(fit)

```


Not all earthquakes have foreshocks and some of thm have aftershocks, finding out which one is which needs much more data than we have now (since we can only use data from the same place on each day) and that's probably why we still can't foretell earthquakes.

## 8


```{r , message=FALSE, warning=FALSE}

depth_intensity_data = worldwide %>% select(mag,depth) %>% na.omit
cor.test(depth_intensity_data$mag, depth_intensity_data$depth)

```
The P-value is very low so we reject H0 and  So we can reject the null hypothesis and conclude that there might be a correlation between depth and magnitude

## 9


```{r , message=FALSE, warning=FALSE}

intensity_mean_data = worldwide %>% filter(type=="earthquake")
intensity_mean_data$time = str_sub(intensity_mean_data$time,0,4)
intensity_mean_data$place = as.character(intensity_mean_data$place)

intensity_mean_data$country = intensity_mean_data$place

places = as.integer(regexpr("\\,[^\\,]*$", intensity_mean_data$country))
intensity_mean_data$country[str_detect(intensity_mean_data$country,",")] = trimws(str_sub(intensity_mean_data$country, places+1))

places = as.integer(regexpr(':', intensity_mean_data$country))
intensity_mean_data$country[str_detect(intensity_mean_data$country,":")] = str_sub(intensity_mean_data$country,1,places-1)

us_states = c("Alabama","Alaska","Arizona","Arkansas","California","Colorado","Connecticut","Delaware","Florida","Georgia","Hawaii","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire","New Jersey","New Mexico","New York","North Carolina","North Dakota","Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island","South Carolina","South Dakota","Tennessee","Texas","Utah","Vermont","Virginia","Washington","West Virginia","Wisconsin","Wyoming")

intensity_mean_data$country[intensity_mean_data$country %in% us_states] = "United States Of America"

intensity_mean_data$country = chartr("a-z", "A-Z",intensity_mean_data$country)

country_intensity = intensity_mean_data %>% group_by(country) %>% summarise(meanMag = mean(mag), count=n())

meanMap = left_join(map, country_intensity) 

hcmap("custom/world-eckert3", 
      data = meanMap, value = "meanMag", 
      joinBy = c("hc-key", "code"), name = "Mean Magnitude for all data", 
      borderColor = "blue", borderWidth = 0.2) %>% 
  hc_mapNavigation(enabled = TRUE) %>% hc_add_theme(hc_theme_google())


country_intensity2 = intensity_mean_data %>% group_by(country,time) %>% summarise(meanMag = mean(mag), count=n()) %>% filter(time==2015)

meanMap2 = left_join(map, country_intensity2) 

hcmap("custom/world-eckert3", 
      data = meanMap2, value = "meanMag", 
      joinBy = c("hc-key", "code"), name = "Mean Magnitude for year 2015", 
      borderColor = "blue", borderWidth = 0.2) %>% 
  hc_mapNavigation(enabled = TRUE) %>% hc_add_theme(hc_theme_google())


country_intensity3 = intensity_mean_data %>% group_by(country,time) %>% summarise(meanMag = mean(mag), count=n()) %>% filter(time==2014)

meanMap3 = left_join(map, country_intensity3) 

hcmap("custom/world-eckert3", 
      data = meanMap3, value = "meanMag", 
      joinBy = c("hc-key", "code"), name = "Mean Magnitude for year 2014", 
      borderColor = "blue", borderWidth = 0.2) %>% 
  hc_mapNavigation(enabled = TRUE) %>% hc_add_theme(hc_theme_google())

```

Looking at the plot we see that many US allies and even US itsself has higher earthquake rates than Iran and Iran doesn't have a particularly high rate in accordance to them, so we can't prove Haarp theory




## 10

find the biggest magnitude for each of the types

```{r , message=FALSE, warning=FALSE}

worldwide %>% group_by(type) %>% summarise(max_mag = max(mag))-> maxes
maxes

```


Find the place with most accidents

```{r , message=FALSE, warning=FALSE}

worldwide %>% group_by(place) %>% summarise(count = n()) %>% arrange(-count) %>% head(10)-> maxes
maxes

```



Number of accidents for each type

```{r , message=FALSE, warning=FALSE}

worldwide %>% group_by(type) %>% summarise(count = n()) %>% arrange(-count) -> nums
nums

```



Number of people killed in each country

```{r , message=FALSE, warning=FALSE}

disaster %>% group_by(COUNTRY) %>% summarise(killed = sum(TOTAL_DEATHS)) %>% arrange(-killed) -> deaths
deaths

```
