---
title: "first week: Introduction to R and Rstudio"
author: "Alaleh Ahmadian Shalchi"
date: "February 2, 2018"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = T)
```

## Problem 1

Create A friendly matrix of size 4,5 and 6.

We could do this by the algorithm (easily searchable on google) to make magic squares
also there's magic function from magic library which generates a magic sguare of given size


```{r , message=FALSE, warning=FALSE}

n = 4
I = matrix(data=0 , n , n)
J = matrix(data=0 , n , 2*n)
tb = matrix(data=0 , n , n)
index=1

for (i in 1:n){
    for (j in 1:n){
      I[i,j]=(i%%4)%/%2;
      J[j,i]=(i%%4)%/%2;
      tb[i,j]=index;
      index = index+1
    }
}

for (i in 1:n){
    for (j in 1:n){
      if (I[i,j]==J[i,j])
        tb[i,j] = n*n+1-tb[i,j];
    }
}

print(tb)


n = 5
nsqr = n*n
i = 1
j = (n+1)/2   
tb = matrix(data=0 , n , n)

for (k in 1:nsqr){

  tb[i,j] = k
  i = i-1
  j = j+1

  if (k%%n == 0){ 
    i = i+2
    j = j-1
  }
  else {
    if (j==n+1){
      j = j - n
    }
    else if (i<1){
      i = i+n
    }
  }
}
print(tb)

# an easier way is using the built-in magic function in magic library

library(magic)
require(abind)
magic(6)


```


## Problem 2

Find the permutations that satisfies all the given constraints

We have 3 groups of different sets of equations
we check to find 5-tuples of numbers 1 to 13 for the first set while removing the tuples that don't satisfy the condition along the way (we could use combn,permn,permutations,... which are all built-in functions of R but using proc.time() function, turns out using optimized while loops is the fastest way to solve the problem)
We do the same thing for the second set of variables (a5 to 10) and then check the two sets for tuples with same a5
In the end we check for the 3 other variables and find the permutations that satisfy all constraints



```{r}

n = 13
perms1to5size = 1
perms1to5 = matrix(data=0,450,5)
perms5to10size = 1
perms5to10 = matrix(data=0,50,6)

a1 = 1
while (a1<=n){
  a2 = 1
  while (a2<=n){
    while (a2<=n && a1==a2){
      a2 = a2+1
    }
    if (a2<=n){
      a3 = 1
      while (a3<=n){
        while(a3<=n && (a1==a3 || a2==a3 || a1>=a2+a3)){
          a3 = a3+1
        }
        if (a3<=n){
          a4 = 1
          while(a4<=n){
            while (a4<=n && (a4==a1 || a4==a2 || a4==a3)){
              a4 = a4+1
            }
            if (a4<=n){
              a5 = 1
              while(a5<=n){
                while (a5<=n && (a5==a1 || a5==a2 || a5==a3 || a5==a4 || a2>=a4+a5 || a1*(a3+a4+a5)!=(a2+a3)*(a2+a3))){
                  a5 = a5+1
                }
                if (a5<=n){
                  	perms1to5[perms1to5size,1]=a1
										perms1to5[perms1to5size,2]=a2
										perms1to5[perms1to5size,3]=a3
										perms1to5[perms1to5size,4]=a4
										perms1to5[perms1to5size,5]=a5
										perms1to5size = perms1to5size+1
                }
              a5 = a5+1
              }
            }
          a4 = a4+1
          }
        }
        a3 = a3+1
      }
    }
    a2 = a2+1
  }
  a1 = a1+1
}

a6 = 1
while (a6<=n){
  a7 = 1
  while (a7<=n){
    while (a7<=n && a6==a7){
      a7 = a7+1
    }
    if (a7<=n){
      a8 = 1
      while (a8<=n){
        while(a8<=n && (a8==a6 || a8==a7)){
          a8 = a8+1
        }
        if (a8<=n){
          a9 = 1
          while(a9<=n){
            while (a9<=n && (a9==a8 || a9==a7 || a9==a6 || a6*(a7+a8+a9)!=(a6+a7)*(a6+a7) || a6>=a8+a9)){
              a9 = a9+1
            }
            if (a9<=n){
              a10 = 1
              while(a10<=n){
                while (a10<=n && (a10==a9 || a10==a8 || a10==a7 || a10==a6)){
                  a10 = a10+1
                }
                if (a10<=n){
                  a5 = 1
                  while (a5<=n){
                    while(a5<=n && (a5==a6 | a5==a7 || a5==a8 || a5==a9 || a5==a10 || a7+a8>=a5+a10 || (a6+a7)*(a5+a9+a10)!=(a7+a8+a9)*(a7+a8+a9) || a6*(a5+a9+a10)!=(a6+a7)*(a7+a8+a9))){
                      a5 = a5+1
                    }
                    if (a5<=n){
                      	perms5to10[perms5to10size,1]=a5
					  		  			perms5to10[perms5to10size,2]=a6
				  		  				perms5to10[perms5to10size,3]=a7
			  		  					perms5to10[perms5to10size,4]=a8
										    perms5to10[perms5to10size,5]=a9
										    perms5to10[perms5to10size,6]=a10
									    	perms5to10size = perms5to10size+1
                    }
                    a5 = a5+1
                  }
                }
              a10 = a10+1
              }
            }
          a9 = a9+1
          }
        }
        a8 = a8+1
      }
    }
    a7 = a7+1
  }
  a6 = a6+1
}

a11 = 1
while (a11<=n){
    a12 = 1
		while (a12<=n){
			while (a12<=n && a11==a12){
				a12 = a12+1
			}
			if (a12<=n){
				for (i in 1:(perms1to5size-1)){
					for (j in 1:(perms5to10size-1)){
						if (perms1to5[i,5]==perms5to10[j,1]){
							a1=perms1to5[i,1]
							a2=perms1to5[i,2]
							a3=perms1to5[i,3]
							a4=perms1to5[i,4]
							a5=perms1to5[i,5]
							a6=perms5to10[j,2]
							a7=perms5to10[j,3]
							a8=perms5to10[j,4]
							a9=perms5to10[j,5]
							a10=perms5to10[j,6]
							a13=91-a1-a2-a3-a4-a5-a6-a7-a8-a9-a10-a11-a12
							if ( a1!=a6 && a1!=a7 && a1!=a8 && a1!=a9 && a1!=a10 && a1!=a11 && a1!=a12 && a1!=a13 &&
								 a2!=a6 && a2!=a7 && a2!=a8 && a2!=a9 && a2!=a10 && a2!=a11 && a2!=a12 && a2!=a13 &&							
								 a3!=a6 && a3!=a7 && a3!=a8 && a3!=a9 && a3!=a10 && a3!=a11 && a3!=a12 && a3!=a13 &&
								 a4!=a6 && a4!=a7 && a4!=a8 && a4!=a9 && a4!=a10 && a4!=a11 && a4!=a12 && a4!=a13 &&
								 a5!=a11 && a5!=a12 && a5!=a13 && a6!=a11 && a6!=a12 && a6!=a13 && 
								 a7!=a11 && a7!=a12 && a7!=a13 && a8!=a11 && a8!=a12 && a8!=a13 &&
								 a9!=a11 && a9!=a12 && a9!=a13 && a10!=a11 && a10!=a12 && a10!=a13 && a12!=a13 && a13<=n && a11<a13 && a12<a10 && (a12+a11)*(a13+a10)==(a12+a13)*(a12+a13)){
								  ans = c(a1,a2,a3,a4,a5,a6,a7,a8,a9,a10,a11,a12,a13)
                  print(ans)
							}
						}
					}
				}
			}
		a12 = a12+1
		}
	a11 = a11+1
}


```

## Problem 3

How many liars can be in each group of people?

for each part we construct all the 0,1-codes, 0 indicating liars and 1 indicating honest people. now we check if there's any paradoxes in each code with the given information of the question, if there's none, we've found a feasible solution


```{r}

mat = matrix(data=0 , 1, 18)
ans = c()

for (i in 1:2^16){
  
  flag = 1
  mat[1,18] = mat[1,2]
  mat[1,1] = mat[1,17]
  for (j in 2:17){
    if ((mat[1,j]==1 && (mat[1,j+1]==1 || mat[1,j-1]==1)) || (mat[1,j]==0 && mat[1,j+1]==0 && mat[1,j-1]==0)){
      flag = 0
    }
  }
  
  if (flag){
    if (!is.element((16-sum(mat[1,2:17])),ans)){
      ans = append(ans,16-sum(mat[1,2:17]))
    }
  }
  
  t = 17
  while(mat[1,t]==1 && t>1){
    t = t-1
  }
  mat[1,t] = 1
  t = t+1
  
  if (t<=17){
    for (j in t:17){
      mat[1,j] = 0
    }
  }
}

print(ans)


mat = matrix(data=0 , 1, 14)
ans = c()

for (i in 1:2^12){
  flag = 1
  mat[1,14] = mat[1,2]
  mat[1,1] = mat[1,13]
  for (j in 2:13){
    if ((mat[1,j]==1 && mat[1,j+1]+mat[1,j-1]!=1) || (mat[1,j]==0 && mat[1,j+1]+mat[1,j-1]==1)){
      flag = 0
    }
  }
  
  if (flag){
    if (!is.element((12-sum(mat[1,2:13])),ans)){
      ans = append(ans,12-sum(mat[1,2:13]))
    }
  }
  
  t = 13
  while(mat[1,t]==1 && t>1){
    t = t-1
  }
  mat[1,t] = 1
  t = t+1
  
  if (t<=13){
    for (j in t:13){
      mat[1,j] = 0
    }
  }
}

print(ans)


mat = matrix(data=0 , 1, 10)
ans = c()
p = combn(2:9,4)

for (i in 1:2^8+1){

  mat[1,10] = mat[1,2]
  mat[1,1] = mat[1,9]
  
  for (k in 1:70){
    flag = 1
    temp = p[,k]
    for (j in 1:4){
      if ((mat[1,temp[j]]==1 && (mat[1,temp[j]+1]+mat[1,temp[j]-1])!=1) ||(mat[1,temp[j]]==0 && (mat[1,temp[j]+1]+mat[1,temp[j]-1])==1)){
        flag = 0
      }
    }
    if(flag){
      for (j in 2:9){
        if (!(is.element(j,temp)) && ((mat[1,j]==1 && (mat[1,j+1]+mat[1,j-1])!=0) ||  (mat[1,j]==0 && (mat[1,j+1]+mat[1,j-1])==0))){
          flag = 0
        }
      }
    }
    
    if(flag==1){
      if (!is.element((8-sum(mat[1,2:9])),ans)){
        ans = append(ans,8-sum(mat[1,2:9]))
      }
    }
  
  }
  
  t = 9
  while(mat[1,t]==1 && t>1){
    t = t-1
  }
  mat[1,t] = 1
  t = t+1
  
  if (t<=9){
    for (j in t:9){
      mat[1,j] = 0
    }
  }
}

print(ans)


```
